---
title: 'Data Science: Wrangling - Assessments'
author: "Hector Antigua"
date: "10/27/2021"
output: word_document

---

```{r,include = FALSE}
options(repos = list(CRAN="http://cran.rstudio.com/"))
```


## Assessment Part 1: Data Import

In this part of the assessment, you will answer several multiple choice
questions that review the concepts of data import. You can answer these
questions without using R, although you may find it helpful to
experiment with commands in your console.

In the second part of the assessment on the next page, you will import
real datasets and learn more about useful arguments to readr functions.
The second part of the assessment will require you to program in R.

### **Question 1**

Which of the following is NOT part of the data wrangling process?

**Respuesta:** *Checking correlations between your variables*

### **Question 2**

Which files could be opened in a basic text editor?

**Respuesta:** *data.txt* *data.csv* *data.tsv*

### **Question 3**

You want to analyze a file containing race finish times for a recent
marathon. You open the file in a basic text editor and see lines that
look like the following:

    initials,state,age,time
    vib,MA,61,6:01
    adc,TX,45,5:45
    kme,CT,50,4:19

**Respuesta:** *A comma-delimited file with a header*

### **Question 4**

Assume the following is the full path to the directory that a student
wants to use as their working directory in R:
"/Users/student/Documents/projects/"

Which of the following lines of code CANNOT set the working directory to
the desired "projects" directory?

**Respuesta:** *setwd(/Users/student/Documents/projects/)*

### **Question 5**

We want to copy the "murders.csv" file from the dslabs package into an
existing folder "data", which is located in our HarvardX-Wrangling
projects folder. We first enter the code below into our RStudio console.

    > getwd()
    [1] "C:/Users/UNIVERSITY/Documents/Analyses/HarvardX-Wrangling"
    > filename <- "murders.csv"
    > path <- system.file("extdata", package = "dslabs")

Which of the following commands would NOT successfully copy
"murders.csv" into the folder "data"?

**Respuesta:** *file.copy(file.path(path, "murders.csv"), getwd())*

### **Question 6**

You are not sure whether the murders.csv file has a header row. How
could you check this?

**Respuesta:** *Open the file in a basic text editor.* *In the RStudio
"Files" pane, click on your file, then select "View File".* *Use the
command read_lines (remembering to specify the number of rows with the
n_max argument).*

### **Question 7**

What is one difference between `read_excel()` and `read_xlsx()`?

**Respuesta:** *`read_excel()` reads both .xls and .xlsx files by
detecting the file format from its extension, while `read_xlsx()` only
reads .xlsx files.*

### **Question 8**

You have a file called "times.txt" that contains race finish times for a
marathon. The first four lines of the file look like this:

    initials,state,age,time
    vib,MA,61,6:01
    adc,TX,45,5:45
    kme,CT,50,4:19

Which line of code will NOT produce a tibble with column names
"initials", "state", "age", and "time"?

**Respuesta:** *race_times \<- read.csv("times.txt")*

### **Question 9**

You also have access to marathon finish times in the form of an Excel
document named "times.xlsx". In the Excel document, different sheets
contain race information for different years. The first sheet is named
"2015", the second is named "2016", and the third is named "2017".

Which line of code will NOT import the data contained in the "2016" tab
of this Excel sheet?

**Respuesta:** *times_2016 \<- read_xlsx("times.xlsx", sheet = "2")*

### **Question 10**

You have a comma-separated values file that contains the initials, home
states, ages, and race finish times for marathon runners. The runners'
initials contain three characters for the runners' first, middle, and
last names (for example, "KME").

You read in the file using the following code.

    race_times <- read.csv(“times.csv”)

What is the data type of the initials in the object `race_times`?

**Respuesta:** *factors* Nota:*In previous versions of R, this was true,
but is not any more. read.csv() no longer automatically converts
characters to factors. If you want to read in character columns as
factors, you can supply the argument stringsAsFactors = T.*

### **Question 11**

Which of the following is NOT a real difference between the readr import
functions and the base R import functions?

**Respuesta:** *The base R import functions can read .csv files, but
cannot read files with other delimiters, such as .tsv files, or
fixed-width files.*

### **Question 12**

You read in a file containing runner information and marathon finish
times using the following code.

    race_times <- read.csv(“times.csv”, stringsAsFactors = F)

What is the class of the object `race_times`?

**Respuesta:** *data frame*

### **Question 13**

Select the answer choice that summarizes all of the actions that the
following lines of code can perform. Please note that the url below is
an example and does not lead to data.

      url <- "https://raw.githubusercontent.com/MyUserName/MyProject/master/MyData.csv "
    dat <- read_csv(url)
    download.file(url, "MyData.csv")

**Respuesta:** *Create a tibble in R called dat that contains the
information contained in the csv file stored on Github. Download the csv
file to the working directory and name the downloaded file
"MyData.csv".*

# Assessment Part 2: Data Import

In this part of the assessment, you will import real datasets and learn
more about useful arguments to readr functions. You will encounter
common issues that arise when importing raw data. This part of the
assessment will require you to program in R.

Use the readr package in the tidyverse library:

```{r}
library(tidyverse)
```

### **Question 14**

Inspect the file at the following URL:

<https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/wdbc.data>

Which **readr** function should be used to import this file?

**Respuesta:** `read_csv()`

### **Question 15**

Check the documentation for the readr function you chose in the previous
question to learn about its arguments. Determine which arguments you
need to the file from the previous question:

```{r}
url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/wdbc.data"
```

Does this file have a header row? Does the readr function you chose need
any additional arguments to import the data correctly?

**Respuesta:** *No, there is no header. The col_names=FALSE argument is necessary.*

### **Question 16**

Inspect the imported data from the previous question.

```{r}
d_q16 <- read_csv(url, col_names = FALSE)

```

How many rows are in the dataset? *569*

How many columns are in the dataset? *32*

# Assessment Part 1: Reshaping Data

Part 1 consists of 8 questions are conceptual questions about tidy data and reshaping data. They do not necessarily require R, but you may benefit from checking your work on the console.

Part 2 consists of 7 questions which require you to write code in R to apply the new concepts about tidy data and reshaping data.

### **Question 1**

A collaborator sends you a file containing data for three years of average race finish times.

```
age_group,2015,2016,2017
20,3:46,3:22,3:50
30,3:50,3:43,4:43
40,4:39,3:49,4:51
50,4:48,4:59,5:01
```
Are these data considered “tidy” in R? Why or why not?

**Respuesta:** *No. These data are not considered “tidy” because the variable “year” is stored in the header.*

### **Question 2**

Below are four versions of the same dataset. Which one is in a tidy format?


**Respuesta:**
```
state      abb region  population total
Alabama     AL	South	4779736	  135
Alaska      AK   West 	710231	  19
Arizona     AZ   West	6392017   232
Arkansas    AR  South	2915918	  93
California  CA   West   37253956  1257
Colorado    CO   West	5029196	  65
```

### **Question 3**

Your file called “times.csv” has age groups and average race finish times for three years of marathons.

```
age_group,2015,2016,2017
20,3:46,3:22,3:50
30,3:50,3:43,4:43
40,4:39,3:49,4:51
50,4:48,4:59,5:01
```
You read in the data file using the following command.

```
d <- read_csv("times.csv")
```

**Respuesta:**    
```
  tidy_data <- d %>%
    gather(year, time, `2015`:`2017`)
```

### **Question 4**

You have a dataset on U.S. contagious diseases, but it is in the following wide format:

```
> head(dat_wide)
state year population HepatitisA Mumps Polio Rubella
Alabama 1990    4040587      86	   19    76    1
Alabama 1991    4066003      39	   14    65    0
Alabama 1992    4097169      35	   12    24    0
Alabama 1993    4133242      40	   22    67    0
Alabama 1994    4173361      72	   12    39    0
Alabama 1995    4216645      75     2    38    0
```

You want to transform this into a tidy dataset, with each row representing an observation of the incidence of each specific disease (as shown below):

```
> head(dat_tidy)
state   year  population  disease  count
Alabama 1990	4040587 HepatitisA	86
Alabama 1991	4066003 HepatitisA	39
Alabama 1992	4097169 HepatitisA	35
Alabama 1993	4133242 HepatitisA	40
Alabama 1994	4173361 HepatitisA	72
Alabama 1995	4216645 HepatitisA	75
```

Which of the following commands would achieve this transformation to tidy the data?

**Respuesta:**   
```
dat_tidy <- dat_wide %>%
    gather(key = disease, value = count, HepatitisA:Rubella)
```

### **Question 5**

You have successfully formatted marathon finish times into a tidy object called tidy_data. The first few lines are shown below.

```
age_group year   time
20        2015   03:46
30        2015   03:50
40        2015   04:39
50        2015   04:48
20        2016   03:22
```

Select the code that converts these data back to the wide format, where each year has a separate column.

**Respuesta:**   
*tidy_data %>% spread(year, time)*

### **Question 6**

You have the following dataset:
```
    > head(dat)
state   abb region    	var   people
Alabama  AL  South population 4779736
Alabama  AL  South  	total 	  135
Alaska   AK   West population  710231
Alaska   AK   West  	total  	   19
Arizona  AZ   West population 6392017
Arizona  AZ   West  	total 	  232
  ```
You would like to transform it into a dataset where population and total are each their own column (shown below):

```
state      abb region population total
Alabama     AL  South	4779736   135
Alaska      AK   West 710231    19
Arizona     AZ   West	6392017   232
Arkansas    AR  South	2915918    93
California  CA   West  37253956  1257
Colorado    CO   West	5029196	   65
```
Which code would best accomplish this?

**Respuesta:**   
*dat_tidy <- dat %>% spread(key = var, value = people)*

### **Question 7**

A collaborator sends you a file containing data for two years of average race finish times, "times.csv":

```
age_group,2015_time,2015_participants,2016_time,2016_participants
20,3:46,54,3:22,62
30,3:50,60,3:43,58
40,4:39,29,3:49,33
50,4:48,10,4:59,14
```
You read in the data file:

```
d <- read_csv("times.csv")

```
Which of the answers below best makes the data tidy?


**Respuesta:**   
```
  tidy_data <- d %>%
    gather(key = “key”, value = “value”, -age_group) %>%
    separate(col = key, into = c(“year”, “variable_name”), sep = “_”) %>% 
    spread(key = variable_name, value = value)
```

### **Question 8**

You are in the process of tidying some data on heights, hand length, and wingspan for basketball players in the draft. Currently, you have the following:

```
    > head(stats)
key               value
allen_height      75
allen_hand_length 8.25
allen_wingspan	  79.25
bamba_height      83.25
bamba_hand_length 9.75
bamba_wingspan    94
 ``` 
Select all of the correct commands below that would turn this data into a “tidy” format with columns "height", "hand_length" and "wingspan".

**Respuesta:**

```
 tidy_data <- stats %>%
    separate(col = key, into = c("player", "variable_name"), sep = "_", extra = "merge") %>% 
    spread(key = variable_name, value = value)
```

# Assessment Part 2: Reshaping Data

Use the following libraries for these questions:


```{r}
library(tidyverse)
library(dslabs)

```

### **Question 9**

Examine the built-in dataset `co2.` This dataset comes with base R, not **dslabs** - just type `co2` to access the dataset.

Is `co2` tidy? Why or why not?

```{r}
head(co2)
```
**Respuesta:**
*co2 is not tidy: to be tidy we would have to wrangle it to have three columns (year, month and value), and then each co2 observation would have a row.*

### **Question 10**

Run the following command to define the co2_wide object:

```{r}
co2_wide <- data.frame(matrix(co2, ncol = 12, byrow = TRUE)) %>% 
      setNames(1:12) %>% 
    mutate(year = as.character(1959:1997))

```

Use the `gather()` function to make this dataset tidy. Call the column with the CO2 measurements `co2` and call the month column `month.` Name the resulting object `co2_tidy.`

Which code would return the correct tidy format?

**Respuesta:**
```{r}
co2_tidy <- gather(co2_wide,month,co2,-year)
```

### **Question 11**

Use co2_tidy to plot CO2 versus month with a different curve for each year:

```{r}
co2_tidy %>% ggplot(aes(as.numeric(month), co2, color = year)) + geom_line()
```

      


    
What can be concluded from this plot?

**Respuesta:**
*CO2 concentrations are highest around May and the yearly average increased from 1959 to 1997.*


### **Question 12**

Load the `admissions` dataset from **dslabs**, which contains college admission information for men and women across six majors, and remove the `applicants` percentage column:

```{r}
library(dslabs)
data(admissions)
dat <- admissions %>% select(-applicants)
```
Your goal is to get the data in the shape that has one row for each major, like this:

```
major  men   women
A      62    82		
B      63    68		
C      37    34		
D      33    35		
E      28    24		
F       6     7	
```
Which command could help you to wrangle the data into the desired format?
      
**Respuesta:**
```{r}
dat_tidy <- spread(dat, gender, admitted)
```

### **Question 13**

Now use the `admissions` dataset to create the object tmp, which has columns `major`, `gender`, `key` and `value`:


```{r}
tmp <- gather(admissions, key, value, admitted:applicants)
tmp
```
Combine the `key` and `gender` and create a new column called `column_name` to get a variable with the following values: `admitted_men`, `admitted_women`, `applicants_men` and `applicants_women.` Save the new data as `tmp2`.

Which command could help you to wrangle the data into the desired format?

**Respuesta:**
```{r}
tmp2 <- unite(tmp, column_name, c(key, gender))
```

### **Question 14**

Which function can reshape `tmp2` to a table with six rows and five columns named `major`, `admitted_men`, `admitted_women`, `applicants_men` and `applicants_women`?

**Respuesta:**
```{r}
spread(tmp2,column_name,value)
```

# Assessment: Combining Tables

### **Question 1**

You have created data frames `tab1` and `tab2` of state population and election data, similar to our module videos:

```
      > tab1
state   	     population
Alabama             4779736
Alaska     	         710231
Arizona    	        6392017
Delaware     	     897934
District of Columbia 601723

> tab2
state  electoral_votes
Alabama      9
Alaska       3
Arizona     11
California  55
Colorado     9
Connecticut  7

> dim(tab1)
[1] 5 2

> dim(tab2)
[1] 6 2
   ``` 
What are the dimensions of the table dat, created by the following command?

      
```dat <- left_join(tab1, tab2, by = “state”)```

**Respuesta:**
*5 rows by 3 columns*

### **Question 2**

We are still using the `tab1` and `tab2` tables shown in question 1. What join command would create a new table “dat” with three rows and two columns?

**Respuesta:**
*dat <- semi_join(tab1, tab2, by = “state”)*

### **Question 3**

Which of the following are real differences between the join and bind functions?

**Respuesta:**
*Binding functions combine by position, while join functions match by variables.*
*Joining functions can join datasets of different dimensions, but the bind functions must match on the appropriate dimension (either same row or column numbers).*
*Bind functions can combine both vectors and dataframes, while join functions work for only for dataframes.*

### **Question 4**

We have two simple tables, shown below, with columns x and y:

```
> df1
 x     y    
 a     a    
 b     a    

> df2
 x     y    
 a     a    
 a     b  
 ```
 
Which command would result in the following table?

```
> final
 x     y    
 b     a   
 ```
 
 **Respuesta:**
 
```
final <- setdiff(df1, df2)
```

**Introduction to Questions 5-7**

Install and load the Lahman library. This library contains a variety of datasets related to US professional baseball. We will use this library for the next few questions and will discuss it more extensively in the Regression course. For now, focus on wrangling the data rather than understanding the statistics.

The Batting data frame contains the offensive statistics for all baseball players over several seasons.  Filter this data frame to define top as the top 10 home run (HR) hitters in 2016:

```{r}
install.packages("Lahman")
library(Lahman)
top <- Batting %>% 
  filter(yearID == 2016) %>%
  arrange(desc(HR)) %>%    # arrange by descending HR count
  slice(1:10)    # take entries 1-10
top %>% as_tibble()

```

Also Inspect the Master data frame, which has demographic information for all players:

```{r}
Master %>% as_tibble()
```
### **Question 5**

Use the correct `join` or `bind` function to create a combined table of the names and statistics of the top 10 home run (HR) hitters for 2016. This table should have the player ID, first name, last name, and number of HR for the top 10 players. Name this data frame top_names.

Identify the join or bind that fills the blank in this code to create the correct table:

```{r}
top_names <- top %>% left_join(Master, by="playerID") %>%
    select(playerID, nameFirst, nameLast, HR)
top_names
```
Which bind or join function fills the blank to generate the correct table?

 **Respuesta:**
 
```
left_join(Master)
```

### **Question 6**

Inspect the `Salaries` data frame. Filter this data frame to the 2016 salaries, then use the correct bind join function to add a salary column to the `top_names` data frame from the previous question. Name the new data frame `top_salary.` Use this code framework:

```{r}
top_salary <- Salaries %>% filter(yearID == 2016) %>%
  right_join(top_names) %>%
  select(nameFirst, nameLast, teamID, HR, salary)
```
Which bind or join function fills the blank to generate the correct table?

**Respuesta:**
 
```
right_join(Master)
```
### **Question 7**

Inspect the `AwardsPlayers` table. Filter awards to include only the year 2016.

How many players from the top 10 home run hitters won at least one award in 2016?

```{r}
AwardsPlayers %>% filter(yearID == 2016) %>% semi_join(top_names, by="playerID") %>%  distinct(playerID) %>% count()
```
How many players won an award in 2016 but were not one of the top 10 home run hitters in 2016?

```{r}
AwardsPlayers %>% filter(yearID == 2016) %>% anti_join(top_names, by="playerID") %>% distinct(playerID) %>% count()

```

# Assessment:  Web Scraping

### **Introduction: Questions 1-3**

Load the following web page, which contains information about Major League Baseball payrolls, into R: [https://web.archive.org/web/20181024132313/http://www.stevetheump.com/Payrolls.htm](https://web.archive.org/web/20181024132313/http://www.stevetheump.com/Payrolls.htm) 

```{r}
library(rvest)
url <- "https://web.archive.org/web/20181024132313/http://www.stevetheump.com/Payrolls.htm"
h <- read_html(url)
```
We learned that tables in html are associated with the table node.  Use the html_nodes() function and the table node type to extract the first table. Store it in an object nodes:

```{r}
nodes <- html_nodes(h, "table")
```
The html_nodes() function returns a list of objects of class xml_node. We can see the content of each one using, for example, the html_text() function. You can see the content for an arbitrarily picked component like this:

```{r}
html_text(nodes[[8]])
```
If the content of this object is an html table, we can use the html_table() function to convert it to a data frame:

```{r}
html_table(nodes[[8]])

```
### **Question 1**

Many tables on this page are team payroll tables, with columns for rank, team, and one or more money values.

Convert the first four tables in nodes to data frames and inspect them. (Note that "parsing errors" and/or "empty tables" still count towards the table index!)

Which of the first four nodes are tables of team payroll?

**Respuesta:**

*Table 2, Table 3 and Table 4*

```{r}
html_table(nodes[1:4])
```

### **Question 2**

For the last 3 components of nodes, which of the following are true? (Check all correct answers.)

**Respuesta:**

*All three entries are tables. *
*The last entry shows the average across all teams through time, not payroll per team.*

```{r}
html_table(tail(nodes,n=3))
```
### **Question 3**

Create a table called `tab_1` using entry 10 of `nodes.` Create a table called `tab_2` using entry 19 of `nodes.`

Note that the column names should be `c("Team", "Payroll", "Average")`. You can see that these column names are actually in the first data row of each table, and that `tab_1` has an extra first column `No.` that should be removed so that the column names for both tables match.

Remove the extra column in `tab_1`, remove the first row of each dataset, and change the column names for each table to `c("Team", "Payroll", "Average")`. Use a `full_join()` by the Team to combine these two tables.

How many rows are in the joined data table?

**Respuesta:** *58*

```{r}
tab_1 <- html_table(nodes[[10]],header = TRUE)[,-1] 
tab_2 <- html_table(nodes[[19]],header = TRUE)
tab_1 %>% full_join(tab_2,"Team")

```
### **Introduction: Questions 4 and 5**

The Wikipedia page on [opinion polling for the Brexit referendum](https://en.wikipedia.org/w/index.php?title=Opinion_polling_for_the_United_Kingdom_European_Union_membership_referendum&oldid=896735054) , in which the United Kingdom voted to leave the European Union in June 2016, contains several tables. One table contains the results of all polls regarding the referendum over 2016:

Use the **rvest** library to read the HTML from this Wikipedia page (make sure to copy both lines of the URL):

```{r}
library(rvest)
library(tidyverse)
url <- "https://en.wikipedia.org/w/index.php?title=Opinion_polling_for_the_United_Kingdom_European_Union_membership_referendum&oldid=896735054"
```

### **Question 4**

Assign `tab` to be the html nodes of the "table" class.

How many tables are in this Wikipedia page?

**Respuesta:** *41*

```{r}
h <- read_html(url)
tab <- html_nodes(h, "table")
tab
```

### **Question 5**

Inspect the first several html tables using `html_table()` with the argument fill=TRUE (you can read about this argument in the documentation). Find the first table that has 9 columns with the first column named "Date(s) conducted".

What is the first table number to have 9 columns where the first column is named "Date(s) conducted"?

**Respuesta:** *6*

```{r, include=FALSE}
html_table(tab[1:9], fill = TRUE)
```


```{r}
html_table(tab[[6]], fill = TRUE)
```


